---
title: "DSM and Heterogeneity modeling"
author: "Jihyeon bae"
date: "2024-01-29"
output: html_document
---


# Introduction

## Background and Goals

- Constructing indicators to identify heterogeneity among an IGO's member states based on their regime types. 
- Estimating the correlation between an IGO's political heterogeneity among members and its institutional design choice of Dispute Settlement Mechanism (DSM).


# Setup and packages

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library("pacman")
p_load(plyr, dplyr, tidyr, ggplot2, tidyverse, RColorBrewer, readxl,
       readr, haven, countrycode,
       rqog, igoR, modelsummary, knitr, kableExtra, flextable, plm)

```

# Loading processed data
```{r, echo=FALSE, warning=FALSE}
igo_analysis<-readRDS("~/Desktop/Regime-Heterogeneity-and-DSM/data/processed/igo_analysis.RDS")
igo_analysis<-igo_analysis%>%filter(year>=1970 & year <=2015)
```


# Data Analysis


## Descriptive Visualization
```{r warning=FALSE}

library(ggplot2)
library(dplyr)

# Optional: reorder IGOs by median DSM for easier reading
igo_analysis <- igo_analysis %>%
  mutate(ioname = reorder(ioname, DSM, median, na.rm = TRUE))

# Violin plot of DSM scores per IGO
p_violin <- ggplot(igo_analysis, aes(x = ioname, y = DSM)) +
  geom_violin(fill = "skyblue", alpha = 0.6, color = "gray30", scale = "width", trim = TRUE) +
  geom_boxplot(width = 0.1, outlier.size = 0.5, alpha = 0.5) +
  labs(
    title = "Distribution of DSM Scores Across IGOs",
    x = "IGO Name",
    y = "DSM Score"
  ) +
  theme_minimal(base_size = 12) +
  theme(
    axis.text.x = element_text(angle = 60, hjust = 1, size = 8),
    plot.title = element_text(face = "bold")
  )

# Save as PNG
ggsave("plot_ioname_distribution.png", p_violin, width = 24, height = 8, dpi = 300)

# OPTIONAL: also create a boxplot only (less visually dense)
p_box <- ggplot(igo_analysis, aes(x = ioname, y = DSM)) +
  geom_boxplot(fill = "lightblue", alpha = 0.7, outlier.size = 0.5) +
  labs(
    title = "Boxplot of DSM Scores per IGO",
    x = "IGO Name",
    y = "DSM Score"
  ) +
  theme_minimal(base_size = 12) +
  theme(
    axis.text.x = element_text(angle = 60, hjust = 1, size = 8),
    plot.title = element_text(face = "bold")
  )

ggsave("plot_ioname_boxplot.png", p_box, width = 24, height = 8, dpi = 300)


```

