---
title: "DSM and Heterogeneity modeling"
author: "Jihyeon bae"
date: "2024-01-29"
output: html_document
---


# Introduction

## Background and Goals

- Constructing indicators to identify heterogeneity among an IGO's member states based on their regime types. 
- Estimating the correlation between an IGO's political heterogeneity among members and its institutional design choice of Dispute Settlement Mechanism (DSM).


# Setup and packages

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library("pacman")
p_load(plyr, dplyr, tidyr, ggplot2, tidyverse, RColorBrewer, readxl,
       readr, haven, countrycode,
       rqog, igoR, modelsummary, knitr, kableExtra, flextable, plm)

```

# Loading Raw data
```{r, echo=FALSE, warning=FALSE}

#Qualities of Government
qog<-rqog::read_qog(which_data = "standard", data_type = "time-series") 

qog<-qog%>%
  filter(year>=1950 & year<2015) 

#Correlates of War
COW<-igoR::igo_year_format3 %>%
  filter(year>=1950)
```

---

# Country-level variables

```{r, echo = FALSE}
country <-qog %>%
  dplyr::select(cname, year, ccodecow, 
         # World Development Indicators
          wdi_gdpcapcon2015, wdi_pop, wdi_trade, 
         # judicial corruption decision
          vdem_jucorrdc, 
          ciri_assn, #freedom of association
          ciri_injud, # independence of the judiciary
        # Security variables
          atop_number, 
        # Democracy Indices 
          vdem_polyarchy, vdem_libdem, vdem_partipdem,
        # Democracy Indices (Alternatives for robustness check)
          p_polity2, bmr_dem, ht_regtype1,
        # Globalization Index (political)
           kofgi_dr_pg=dr_pg,
        # colonial legacy
           ht_colonial,
        # Political Terror Scale
         pts_ptsa=gd_ptsa) %>%  
  dplyr::mutate(wdi_log_gdpcapcon2015=log(wdi_gdpcapcon2015),
         wdi_log_pop=log(wdi_pop),
         wdi_log_trade=log(wdi_trade))%>%
  dplyr::rename(ccode=ccodecow)%>%
  dplyr::relocate(ccode, cname, year)

country <- country %>%
  mutate(ht_colonial=ifelse(ht_colonial>1, 1, 0),
         ht_regtype1=case_when(ht_regtype1==9 ~ 5, 
                            ht_regtype1==99 ~ 6,
                            ht_regtype1==100 ~ 7,
                            TRUE ~ ht_regtype1))

```

```{r, echo = FALSE}
#Check rows that have NA values for VARIABLE_OF_INTEREST
country[is.na(country$ccode),]

#Fill in NAs
country$ccode[country$cname == "Ethiopia"] <- 530
country$ccode[country$cname == "Germany"] <- 255
country$ccode[country$cname == "Yemen Democratic"] <- 680
country$ccode[country$cname == "Yemen"] <- 679
country$ccode[country$cname == "Sudan"] <- 625
country$ccode[country$cname == "Cyprus"] <- 352
country$ccode[country$cname == "Pakistan"] <- 770
country$ccode[country$cname == "Vietnam, North"] <- 816
country$ccode[country$cname == "Vietnam, South"] <- 817
country$cname[country$cname == "USSR"] <- "Russian Federation (the)"


#Run below code to check year-availability for VARIABLE_OF_INTEREST
#summary(country[!is.na(country$VARIABLE_OF_INTEREST),])

country<-country%>%
  filter(!is.na(ccode))
```

---

## Transposing IGO-level data into country-level data

- Pivot longer so that row represents country
```{r}
COW <- COW %>%
  dplyr::rename(cow_igocode = ionum)%>%
  dplyr::select(-c(igocode, version, accuracyofpre1965membershipdates,sourcesandnotes, imputed)) %>%
  dplyr::relocate(cow_igocode, ioname, year, political, social, economic)%>%
  pivot_longer(c(`afghanistan`:`zimbabwe`),
                      names_to="country",
                      values_to="membership")%>%
  dplyr::filter(membership==1) #member states only

```

## Identify ambiguous country names
```{r}
COW <- COW %>%
  dplyr::mutate(country = dplyr::case_when(
    country == "austriahungary" ~ "Austria-Hungary",
    country == "domrepublic" ~ "Dominican Republic",
    country == "etimor" ~ "East Timor",
    country == "hessegrand" ~ "Hesse Grand Ducal",
    country == "micronesiafs" ~ "Federated States of Micronesia",
    country == "nokorea" ~ "North Korea",
    country == "soafrica" ~ "South Africa",
    country == "sokorea" ~ "South Korea",
    country == "stlucia" ~ "St. Lucia",
    country == "wgermany" ~ "German Federal Republic",
    country == "syemen" ~ "Yemen People's Republic",
    TRUE ~ country  
    # keep original value for all other cases
  ))

#Attaching Country numeric code to character values
COW$ccode<-countrycode(COW$country, 
              origin='country.name', 
              destination='cown', 
              warn = TRUE)

COW<-COW%>%
  filter(!is.na(ccode))
```

```{r}
# accession data
join_events <- COW %>%
  dplyr::group_by(ccode, ioname) %>%
  dplyr::summarise(join_year = min(year, na.rm = TRUE), .groups = "drop") %>%
  dplyr::left_join(country %>% select(ccode, cname) %>% distinct(), by = "ccode")

join_counts <- join_events %>%
  dplyr::group_by(cname, join_year) %>%
  dplyr::summarise(n_igos = n(), .groups = "drop")

country_ts <- country %>%
  select(ccode, cname, year, vdem_polyarchy)

selected_countries <- c("United States", "Germany", "India", "Brazil", "China", "Korea (the Republic of)")

country_ts_subset <- country_ts %>%
  filter(cname %in% selected_countries)

join_events <- join_events %>%
  filter(cname %in% selected_countries)

join_counts <- join_counts %>%
  filter(cname %in% selected_countries)
library(ggplot2)

ggplot() +
  # Democracy line
  geom_line(
    data = country_ts_subset,
    aes(x = year, y = vdem_polyarchy),
    color = "steelblue"
  ) +
  
  # Points showing counts of IGOs joined
  geom_point(
    data = join_counts,
    aes(
      x = join_year,
      y = 0,
      size = n_igos
    ),
    color = "red",
    alpha = 0.7
  ) +
  ylim(0,1) +
  facet_wrap(~ cname, scales = "free_y") +
  labs(
    title = "Country Democracy Trajectories with IGO Accession Events",
    subtitle = "Red points = Number of IGOs joined",
    x = "Year",
    y = "V-Dem Polyarchy Score",
    size = "Number of IGOs joined"
  ) +
  theme_minimal(base_size = 12) +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    strip.text = element_text(face = "bold")
  )

```

## missing
```{r}
missing <- ggplot(country_ts, aes(x = year, y = cname, color = is.na(vdem_polyarchy))) +
  geom_point(size = 1) +
  scale_color_manual(
    values = c("black", "red"),
    labels = c("Present", "Missing"),
    name = "Polyarchy"
  ) +
  labs(
    title = "Missing V-Dem Polyarchy Scores by Country-Year",
    x = "Year",
    y = "Country"
  ) +
  theme_minimal(base_size = 8) +
  theme(
    axis.text.x = element_text(
      angle = 90,
      vjust = 0.5,
      hjust = 1,
      size = 5
    ),
    plot.margin = margin(10, 10, 20, 10)
  ) +
  coord_flip()



ggsave("missing.png", missing, width = 30, height = 8, dpi = 500)

```



## Extracting member-state information for each IGO
```{r}
igo_master <- COW %>%
  dplyr::inner_join(country, by=c("ccode", "year"))%>%
  dplyr::select(-c(orgname, longorgname, membership))


igo_master <- igo_master %>%
  group_by(cow_igocode, year) %>%
    mutate(gdp_share = 100*(wdi_gdpcapcon2015/sum(wdi_gdpcapcon2015, na.rm = TRUE)),
    poly_share = 100*(vdem_polyarchy/sum(vdem_polyarchy, na.rm = TRUE)))

```



# Variable Generation
## summarising state level info to igo level
```{r}
igo <- igo_master %>%
  dplyr::group_by(cow_igocode, ioname, year) %>%
  dplyr::summarise(
    #average democracy scores
    polyarchy = mean(vdem_polyarchy, na.rm = TRUE),
    w.polyarchy = weighted.mean(vdem_polyarchy, w = wdi_log_gdpcapcon2015/sum(wdi_log_gdpcapcon2015), na.rm=TRUE),
    polyarchy_median = median(vdem_polyarchy, na.rm = TRUE),
    partipdem = mean(vdem_partipdem, na.rm=TRUE),
    partipdem_median = median(vdem_partipdem, na.rm=TRUE),
    libdem = mean(vdem_libdem, na.rm = TRUE),
    libdem_median = median(vdem_libdem, na.rm = TRUE),
    polity = mean(p_polity2, na.rm = TRUE),
    #economic variables
    gdp_cap = mean(wdi_log_gdpcapcon2015, na.rm = TRUE),
    population=mean(wdi_log_pop, na.rm=TRUE),
    trade=mean(wdi_log_trade, na.rm=TRUE),
    alliances = mean(atop_number, na.rm = TRUE),
    # Governance
    ciri_injud=mean(ciri_injud, na.rm=TRUE),
    #characteristics 
    number = n(),
    trade = mean(wdi_log_trade, na.rm = TRUE),
    percentage = sum(bmr_dem, na.rm = TRUE) / number,
    political = mean(political, na.rm = TRUE),
    social = mean(social, na.rm = TRUE),
    economic = mean(economic, na.rm = TRUE),
    colonial= sum(ht_colonial, na.rm=TRUE) / number,
    #asymmetry index
    polity_sd = sd(p_polity2, na.rm = TRUE),
    polyarchy_sd = sd(vdem_polyarchy, na.rm = TRUE),
    libdem_sd=sd(vdem_libdem, na.rm=TRUE),
    partipdem_sd=sd(vdem_partipdem, na.rm=TRUE),
    econ_sd = sd(wdi_log_gdpcapcon2015, na.rm = TRUE),
    # HH index
    hh_poly = sum(abs(poly_share), na.rm = TRUE) - (1 / n()),
    hh_gdp = sum(abs(gdp_share), na.rm = TRUE) - (1 / n()),
    # Coefficient of variation
    polity_cv = polity_sd / polity,
    polyarchy_cv = polyarchy_sd / polyarchy,
)

  
```

