The goal of this script is to compare sample mean and theoretical mean when parameters of beta distribution change. 
Assuming that an IGO is composed of member states whose democracy scores are assigned from beta distribution, this comparison should tell us about the significance of falsely relying on sample mean. 
I hypothesize that the disparity between the two should be more salient *especially* when the data is skewed and when member states are few. 

# inital code for four cases alone
```{r}
set.seed(2025)

params <- list(
  "alpha >1, beta >1" = c(alpha=5, beta=5),
  "alpha <1, beta <1" = c(alpha=0.5, beta=0.5),
  "alpha >1, beta <1" = c(alpha=2, beta=0.5),
  "alpha <1, beta >1" = c(alpha=0.1, beta=5)
)


results <- lapply(names(params), function(name) {
  a <- params[[name]]["alpha"]
  b <- params[[name]]["beta"]
  samp <- rbeta(50, shape1 = a, shape2 = b)
  tibble(
    case = name,
    alpha = a,
    beta = b,
    theoretical_mean = a / (a + b),
    sample_mean = mean(samp),
    theoretical_variance = (a * b) / ((a + b)^2 * (a + b + 1)),
    sample_variance = var(samp)
  )
})

knitr::kable(dplyr::bind_rows(results), digits = 4)

```




I expand from the first two exercises. 

1) General function that compares the two values depending on parameters as inputs. 
```{r}
simulate_beta <- function(alpha, beta, n) {
  parameters <- rbeta(n=n, shape1 = alpha, shape2 = beta)
  
  tibble::tibble(
    alpha = alpha,
    beta = beta,
    n = n,
    theoretical_mean = alpha / (alpha + beta),
    sample_mean = mean(parameters),
    theoretical_variance = (alpha * beta) / (((alpha + beta)^2) * (alpha + beta + 1)),
    sample_variance = var(parameters)
  )
}

simulate_beta(alpha = 3, beta = 2, n = 100)

  
```
With this function, I'm generating a list of two vetors.E(X) vs. theoretical mean.
By simulating 1000 beta-distributed samples of size n=50 for a given a, b pair.

# Case 1) a > 1, b>1 
# should expect a unimodal distrubion
```{r}
reps = 1000
sample_means <- numeric(reps)
sample_vars <- numeric(reps)
theoretical_means <- numeric(reps)
theoretical_vars <- numeric(reps)


for (i in 1:reps) {
  result <- simulate_beta(alpha=5, beta=2, n=50)
  sample_means[i] <- result$sample_mean
  sample_vars[i] <- result$sample_variance
  theoretical_means[i] <- result$theoretical_mean
  theoretical_vars[i] <- result$theoretical_variance
}

sim1 <- tibble::tibble(
  sample_mean = sample_means,
  sample_variance = sample_vars,
  theoretical_mean = theoretical_means,
  theoretical_variance = theoretical_vars
) 
```

# Case 2) a < 1, b < 1
# convex bimodal 
# a group of full autocracies & full democracies
```{r}
reps = 1000
sample_means <- numeric(reps)
sample_vars <- numeric(reps)
theoretical_means <- numeric(reps)
theoretical_vars <- numeric(reps)


for (i in 1:reps) {
  result <- simulate_beta(alpha=0.5, beta=0.5, n=50)
  sample_means[i] <- result$sample_mean
  sample_vars[i] <- result$sample_variance
  theoretical_means[i] <- result$theoretical_mean
  theoretical_vars[i] <- result$theoretical_variance
}

sim2 <- tibble::tibble(
  sample_mean = sample_means,
  sample_variance = sample_vars,
  theoretical_mean = theoretical_means,
  theoretical_variance = theoretical_vars
) 
```

# case 3) a > 1, b < 1
# group of democracies forming the majority
```{r}
reps = 1000
sample_means <- numeric(reps)
sample_vars <- numeric(reps)
theoretical_means <- numeric(reps)
theoretical_vars <- numeric(reps)


for (i in 1:reps) {
  result <- simulate_beta(alpha=5, beta=0.5, n=50)
  sample_means[i] <- result$sample_mean
  sample_vars[i] <- result$sample_variance
  theoretical_means[i] <- result$theoretical_mean
  theoretical_vars[i] <- result$theoretical_variance
}

sim3 <- tibble::tibble(
  sample_mean = sample_means,
  sample_variance = sample_vars,
  theoretical_mean = theoretical_means,
  theoretical_variance = theoretical_vars
) 
```


# case 4) a < 1, b > 1
# group of autocracies forming the majority (*AIGO*)

```{r}
reps = 1000
sample_means <- numeric(reps)
sample_vars <- numeric(reps)
theoretical_means <- numeric(reps)
theoretical_vars <- numeric(reps)


for (i in 1:reps) {
  result <- simulate_beta(alpha=0.5, beta=5, n=50)
  sample_means[i] <- result$sample_mean
  sample_vars[i] <- result$sample_variance
  theoretical_means[i] <- result$theoretical_mean
  theoretical_vars[i] <- result$theoretical_variance
}

sim4 <- tibble::tibble(
  sample_mean = sample_means,
  sample_variance = sample_vars,
  theoretical_mean = theoretical_means,
  theoretical_variance = theoretical_vars
) 
```


2) Tracing the gap between theoretical and sample mean for one IGO over time. 
```{r}
library(fitdistrplus)
# example with just one IGO 
subset <- igo_master %>%
  filter(ioname == "ASEAN") %>%
  pull(v2x_polyarchy)

subset <- subset[!is.na(subset)]

years <- 1950:2014
asean_results <- list()

for (yr in years) {
  subset <- igo_master %>%
    filter(ioname == "ASEAN", year == yr, !is.na(v2x_polyarchy)) %>%
    pull(v2x_polyarchy)
  
  if (length(subset) <= 1) next
  
  fit <- fitdist(subset, "beta", method = "mle")
  
  alpha <- fit$estimate["shape1"]
  beta <- fit$estimate["shape2"]
  
  sim_result <- simulate_beta(alpha = alpha, beta = beta, n = length(subset)) 
  # n varies over time. with new members. 
  
  sim_result$year <- yr
  sim_result$actual_sample_mean <- mean(subset)
  sim_result$actual_sample_variance <- var(subset)
  sim_result$n_actual <- length(subset)
  
  asean_results[[as.character(yr)]] <- sim_result
}

asean_df <- bind_rows(asean_results) 
#n_actual = observations without any NAs.

```

